New Insights on the Role of Context in the Perception of a Human-like Robot

### 2015-04-21 main.tex
Reworked the related work

@@

To study this effect, we propose an original methodology based on eye-tracking
and a serie of stimuli that are visually identical and yet elicit different
assumptions regarding the cognitive capabilities of the robot. We then correlate
the gaze patterns with established questionnaires that assess anthropomorphic
attributions.

@@

Measuring anthropomorphism raises interesting methodological challenges, if only
tbecause people's perception of robots change over the duration of their
interaction with the robots. Takayama discusses the effects of this
change~\cite{takayama_perspectives_2012} and differentiates between what she calls an
\emph{in-the-moment} and a \emph{reflective} perspective on agency: an
\emph{in-the-moment} perspective would refer to one's most immediate
response/sense in a given situation, while a \emph{reflective}
perspective would indicate one's reaction/sense of a situation based on a more
in-depth  consideration and contemplation.

Quite often, these two distinctions are not considered differently, leading to
errors and confusions.  In other words, in an initial phase of interaction with
a non-human agent, people might respond ``mindlessly'' instead of responding
consciously~\cite{nass_machines_2000}. Only post a due amount of
time of, what is generally called as ``familiarization'' with the robot, one
might respond in a more reflective manner instead of an in-the-moment reaction.
This can be demonstrated by the fact that when participants who had interacted
with a technology, were asked about whether the interaction with the technology
was in a human-like way, most of them denied this which shows the reflective
perspective . However, the same participants had actually reacted ( in the
moment of the interaction) to the system in many ways that were quite close to
how they would react to people~\cite{reeves_media_1996}.

Most of the experiments that investigate anthropomorphism in HRI are conducted
in controlled lab-settings during a short-term scenario. For instance, pictures
or videos of different robots are shown to human subjects who subsequently fill
in a questionnaire, to assess how far the different types of robots were
perceived human-like. Such experimental settings have limitations for studying a
social phenomenon like anthropomorphism, and one may wonder if anthropomorphism
can really be measured in a post-interaction closed questionnaire or by rating
scales. Findings might possibly lead to interpretations that would not be
generalizable to natural settings or real interaction scenarios (which is always
an issue in social science research).

The Godspeed questionnaire~\citep{bartneck_measurement_2008}, and in particular,
its subpart focused on anthropomorphism, is the main validated questionnaire to
assess anthropomorphism. On 5 point semantic differential scales, people are
asked to rate the following constructs: fake \vs natural, machinelike \vs
humanlike, unconscious \vs conscious, artificial \vs lifelike, moving rigidly
\vs moving elegantly. Because the concept of ``human-likeness'' itself is
complex and abstract, \cite{kahn_jr._robotic_2006} suggest to ask for more
concrete constructs that are typical or unique of the concept of
``human-likeness'', and \cite{ruijten_introducing_2014} propose for instance a
25-item questionnaire to measure various concrete aspects of human-likeness.

@@

@@

@@

>>>>

\newcommand{\h}[1]{\textbf{H#1}\xspace}

@@

To study this effect, we propose an original methodology based on eye-tracking
and a set of stimuli that are visually identical and yet elicit different
assumptions regarding the cognitive capabilities of the robot. We then correlate
the gaze patterns with established questionnaires that assess anthropomorphic
attributions.

@@

\paragraph{Measuring anthropomorphism} Measuring anthropomorphism raises
interesting methodological challenges, if only because people's perception of
robots change over the duration of their interaction with the
robots~\cite{lemaignan2014dynamics}. Takayama discusses the effects of these
changes~\cite{takayama_perspectives_2012} and differentiates between what she
calls an \emph{in-the-moment} and a \emph{reflective} perspective on agency: an
\emph{in-the-moment} perspective would refer to one's most immediate
response/sense in a given situation, while a \emph{reflective} perspective would
indicate one's reaction/sense of a situation based on a more in-depth
consideration and contemplation.  In the first phase of interaction with a
non-human agent (a robot in our case), people might respond ``mindlessly''
instead of responding consciously~\cite{nass_machines_2000}, and only after what
is called the \emph{familiarization} period, one might respond in a more
reflective manner instead of an in-the-moment reaction: Reeves
\etal~\cite{reeves_media_1996} showed for instance that subjects interacting
with technologies in ways close to how they would behave towards people, would
deny any human-like interaction in a post-hoc questionnaire, illustrating this
reflective perspective.

Yet, most of the experiments that investigate anthropomorphism in HRI are based
on post-interaction, closed questionnaires or rating scales.  The Godspeed
questionnaire~\cite{bartneck_measurement_2008}, and in particular, its subpart
focused on anthropomorphism, is the main validated questionnaire to assess
anthropomorphism. On 5 point semantic differential scales, people are asked to
rate the following constructs: fake \vs natural, machinelike \vs humanlike,
unconscious \vs conscious, artificial \vs lifelike, moving rigidly \vs moving
elegantly. Because the concept of ``human-likeness'' itself is complex and
abstract, Kahn \etal~\cite{kahn_jr._robotic_2006} suggest to ask for more
concrete constructs that are typical or unique of the concept of
``human-likeness'', and Ruijten \etal~\cite{ruijten_introducing_2014} propose
for instance a 25-item questionnaire to measure various concrete aspects of
human-likeness. Other researchers~\cite{zlotowski2014dimensions,salem2015would}
also applied a two-dimensional scale measuring \emph{Human Nature} and
\emph{Uniquely Human} traits of robots, based on an original proposal by Haslam
\etal~\cite{haslam2008attributing}.

We take here a different approach to measure anthropomorphism, based on
eye-tracking.  By looking at the eyes fixation on the head of the robot, we have
been able to evidence significantly different patterns in two experimental
conditions subtly different in terms of human-likeness. This eye-tracking based
metric is behavioural and hence does not suffer post-hoc reconstruction by the
subjects, possibly providing a better, ``noise-less'', access to the human
attitude towards the machine, and, as such, a better proxy to the cognitive and
affective bonds that the human establishes with the robot.

@@

The study has been designed out of the following intuition: \emph{the more we
ascribe cognitive capabilities to an agent, the more we focus on its head when
interacting with it}.  Eye-tracking was a natural methodology to investigate
this intuition, and we crafted audio-visual stimuli that would elicit different
levels of cognitive ascription onto the robot while being identical in every
other way.

Several hypotheses ensued: \h{1}

@@

%%%%%%%%%%%%%%%%%%%%%%

No correlation between delta opinion and gaze fixation on head: are we measuring
two different things? or hidden variable?

If different things, then, gaze is possibly a better metric: in-the-moment, does
not suffer post-hoc reconstruction.

%%%%%%%%%%%%%%%%%%%%%%

@@

Anthropomorphism as a proxy for cognitive/affective bonds that are esthablishing
between humans and robots.

\fxnote{discuss extensions towards mobile eyetrackers}
